---
title: "Introduction to coxdiag"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to coxdiag}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Overview

The `coxdiag` package automates the diagnostic workflow for Cox proportional
hazards models. Given a dataset and a model formula, it fits the model, tests
the proportional hazards assumption, and generates diagnostic plots.

## Installation
```{r eval=FALSE}
devtools::install_github("douyadou0304-dev/coxdiag")
```

## Example: Lung Cancer Dataset

We demonstrate the package using the `lung` dataset from the `survival` package,
which contains survival data from 228 patients with advanced lung cancer.
```{r}
library(coxdiag)
library(survival)

# Fit model and run diagnostics
result <- run_diagnostics(lung, Surv(time, status) ~ age + sex + ph.ecog)
```

## Model Results
```{r}
print(result$model)
```

The model shows that `sex` and `ph.ecog` are significant predictors of survival,
while `age` is not significant after adjusting for the other covariates.

## Proportional Hazards Assumption
```{r}
print(result$ph_test)
```

All p-values are well above 0.05, indicating that the proportional hazards
assumption is not violated for any variable. The global test (p = 0.22) also
supports this conclusion.

## Diagnostic Plots

### Schoenfeld Residuals
```{r fig.width=7, fig.height=6}
print(result$ph_plot)
```

The Schoenfeld residual plots show no systematic trend over time for any variable,
consistent with the formal test results above.

### Martingale Residuals
```{r fig.width=6, fig.height=5}
print(result$martingale_plot)
```

The Martingale residual plot shows no strong nonlinear pattern, suggesting that
the linear functional form for continuous covariates is reasonable.

## Human-Readable Summary
```{r}
print_summary(result)
```

## Conclusion

The `coxdiag` package provides a one-function workflow for Cox model diagnostics.
For this dataset, the model assumptions appear to be satisfied, and the results
suggest that sex and ECOG performance status are the primary predictors of
survival in this lung cancer cohort.

